---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
<style type="text/css">
.table {

    width: 40%;

}
</style>

## Loading and preprocessing the data  
Read the data from csv  
```{r}
activity <- read.csv("activity.csv")
```
Convert date column from character to Date
```{r}
activity$date <- as.Date(activity$date)
```

## Steps taken per day

Get total steps per day and a histogram of the steps variable
```{r}
library(dplyr)
sum_by_day <- activity %>% group_by(date) %>% summarise_at(vars(1), funs(sum(., na.rm=TRUE)))
hist(sum_by_day$steps)
```

Get mean and median  
```{r}
mean_s <- mean(sum_by_day$steps)
median_s <- median(sum_by_day$steps)
```

Mean steps per day : `r formatC(mean_s, digits = 1, format = "f")`  
Median steps per day :  `r median_s`

## Average daily activity pattern

Get the mean steps by interval  
```{r}
library(ggplot2)
library(scales)
mean_by_interval <- activity %>% group_by(interval) %>% summarise_at(vars(1), funs(mean(., na.rm=TRUE)))  
```

Add a time column of class POSIXct
```{r}
mean_by_interval$time <- as.POSIXct( "1970-01-01", tz = "CET" ) + as.difftime( floor(mean_by_interval$interval/100)*60+mean_by_interval$interval %% 100, units="mins"
)
```
Note: 1970 date is not important we will just use the time of day  


```{r}
ggplot(mean_by_interval, aes(x=time, y=steps)) +
  geom_line() +
  scale_x_datetime(name = "Time", date_labels = "%H:%M")+
  scale_y_continuous(name="Steps")+
  ggtitle("Average Daily Activity Pattern") +
  theme(plot.title = element_text(hjust = 0.5))

the_max <- mean_by_interval[which.max(mean_by_interval$steps),]


```

The 5 minutes interval which on average has more steps is `r the_max$interval` with `r the_max$steps` steps average, and corresponds to `r format(the_max$time, "%H:%M")` in the morning

## Imputing missing values

Get number of NAs   
```{r}
nas <- sum(is.na(activity$steps))
```

Number of NA values : `r nas`  


Get a dataframe with NAs filled with the steps mean for that 5 minutes interval
```{r}
activity_p <- activity

the_na_rows <- which(is.na(activity_p$steps))
for (i in the_na_rows){
  its_interval <- activity_p[i,]$interval
  activity_p[i,]$steps <- mean_by_interval[mean_by_interval$interval==its_interval,]$steps
}
```


Get the sum per day for the created dataframe, and a histogram of steps variable  
```{r}
sum_by_day_p <- activity_p %>% group_by(date) %>% summarise_at(vars(1), funs(sum(., na.rm=TRUE)))
hist(sum_by_day_p$steps)
```

Get the mean and the median of created dataframe  
```{r}
mean_s_p <- mean(sum_by_day_p$steps)
median_s_p <- median(sum_by_day_p$steps)
```

Compare raw data mean and median with filled up data mean and median  

-- | Mean | Median |
-- | -- | -- |
Raw | `r formatC(mean_s, digits = 1, format = "f") ` | `r formatC(median_s, digits = 0, format = "f")` |
Filling NAs | `r formatC(mean_s_p, digits = 1, format = "f")` | `r formatC(median_s_p, digits = 0, format = "f") ` |



## Differences in activity patterns between weekdays and weekends

Add factor column to differentiate weekends from weekdays  
```{r}
weekend_days <- c("Saturday", "Sunday")

activity_p$day_type <- as.factor(ifelse(weekdays(activity_p$date) %in% weekend_days, "weekend", "weekday"))
```

Get the mean by interval  
```{r}
mean_by_interval_p <- activity_p %>% group_by(interval, day_type) %>% summarise_at(vars(1), funs(mean(., na.rm=TRUE)))
```

Add time column to construct the plot  
```{r}
mean_by_interval_p$time <- as.POSIXct( "1970-01-01", tz = "CET" ) + as.difftime( floor(mean_by_interval_p$interval/100)*60+mean_by_interval_p$interval %% 100, units="mins"
)
```
Note: 1970 date is not important as we will just use the time of the day  

Plot the differences in activity patern between weekdays and weekends
```{r}
ggplot(data = mean_by_interval_p, mapping = aes(x = time, y = steps)) +
    geom_line() +
    scale_x_datetime(name = "Time", date_labels = "%H:%M")+
    facet_grid(day_type ~ .)

```
